---
title: "Supermarket Data EDA"
author: "William Lovejoy"
date: "6/13/2022"
output: html_document
---

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(reshape2)
```

Quick and easy analysis of supermarket data pulled from [link](https://www.kaggle.com/datasets/surajjha101/stores-area-and-sales-data).

```{r import}
df <- read.csv("C:\\Users\\William Lovejoy\\Documents\\Codes\\R\\DataScience\\SupermarketData\\Stores.csv")
glimpse(df)
```

```{r fig1}
plot(df[, -c(1)])
```
Right away there is a strong correlation between a stores area and the number of
items it has available. 

```{r expansion}
df$Monthly_Customer_Count <- df$Daily_Customer_Count * 30
df$Average_Daily_Sales <- df$Store_Sales / 30
```


After breaking down the data a bit more to see how much a store tends to make
each month, as well as how many customers they tend to receive monthly.

Next we can calculate the actual correlative values for the different variables
and plot them with a tile graph.
```{r fig2}
correlated <- df[, -c(1)] %>%
  cor() %>%
  melt()

ggplot(correlated, aes(x = Var1, y = Var2, fill = value)) + 
  scale_fill_gradient2(low = "antiquewhite",
                       mid = "gold",
                       high = "firebrick") +
  geom_tile(color = "black") +
  geom_text(aes(label = round(value, 2)), size = 5) +
  theme(axis.text.x = element_text(angle = 90), axis.title = element_blank()) +
  labs(title = "Correlation of Supermarket Factors")
```


